---
title: "Lab 0"
author: "Mitchell Harrison"
output: pdf
---

## Exercise 1
```{r}
x = 0.1
y = 0.1
z = 1
(x + y) + z == x + (y + z)
```

The above cell returns `FALSE`, even though the answers should be identical.

```{r}
all.equal((x + y) + z, x + (y + z))
```

Using the `all.equal()` function, we see that the arithmetic from the first
code chunk are considered "equal" after the `all.equal()` function corrects for
some floating point arithmetic peculiarities.

## Exercise 2
```{r}
rnorm(10, mean = 1, sd = 2)
pnorm(0)
dnorm(0.5)
qnorm(0.5)
```

The `rnorm()` outputs an array (here of length 10) with a given mean and
standard deviation (here 1 and 2 respectively).

The `dnorm()` function calculates the value of the PDF of a normal distribution
at a given $x$ value. The following function will calculate it manually:
$$
f(x) = \frac{1}{\sqrt{2\pi}}e^{-\frac{0.5}{2}} \approx 0.3520653
$$

## Exercise 3
```{r}
X <- rnorm(100, mean = 0, sd = 1)
Y <- rnorm(100, mean = 1, sd = sqrt(2))
mean(X + Y)
var(X + Y)
```

